---
output:
 html_document:
  toc: true
  toc_depth: 1
  toc_float:
   collapsed: false
---

```{r, child = 'cdtcsschunks.Rmd'}
```

<style type="text/css">
div.tocify {
 max-height: 60%;
}
</style>

<div id='toc-index' class='toc-index-class'>

# [CDT GUI](cdt_file_menu.html)

## [Starting CDT](cdt_file_menu.html#cdt.starting)

## [CDT Main Window](cdt_file_menu.html#cdt.main.win)

### [Menu Bar](cdt_file_menu.html#menu.bar)

### [Toolbars](cdt_file_menu.html#toolbars)

### [Open Files](cdt_file_menu.html#open.files)

### [Input and Command Panels](cdt_file_menu.html#cmd.panel)

### [Map and Table View](cdt_file_menu.html#map.view)

### [Output Message](cdt_file_menu.html#out.msg)

### [Status Bar](cdt_file_menu.html#status.bar)

# [CDT Input Data Format](cdt_data_input.html)

## [CDT stations data format](cdt_data_input.html#cdt.format.data)

## [Setting date range](cdt_data_input.html#cdt.date.range)

## [CDT coordinates file](cdt_data_input.html#cdt.coords.data)

## [NetCDF data](cdt_data_input.html#cdt.ncdf.data)

## [NetCDF filename format](cdt_data_input.html#cdt.ncdf.ffrmt)

## [Setting NetCDF data](cdt_data_input.html#cdt.ncdf.setting)

# [Format CDT Input data](format_cdt_data.html)

## [Multiple Files](format_cdt_data.html#multiple.files)

## [Single File](format_cdt_data.html#single.file)

# [Merge two CDT Stations Data](merge_2cdt_data.html)

# [Filtering CDT Station Data](filter_cdt_data.html)

# [Transform NetCDF files into CDT Dataset](transform_cdt_dataset.html)

# [Download Data](download_data.html)

## [Download DEM](download_data.html#download.dem)

## [Download Country boundary](download_data.html#download.shp)

## [Download RFE](download_data.html#download.rfe)

# [Filling missing dekadal temperature values](fill_na_temp.html)

# [Data Format Conversion](data_conversion.html)

## [Conversion to CPT data format](data_conversion.html#conversion.cpt)

## [Converting NetCDF/GeoTIFF/ESRI .hdr Labelled](data_conversion.html#conversion.raster)

## [Create GrADS Descriptor File from CDT NetCDF files](data_conversion.html#conversion.grads)

# [Aggregating Time Series](aggregate_TS.html)

# [Aggregate/Disaggregate Spatial NetCDF data](aggregate_SP.html)

# [Check Stations Coordinates](check_stations_coords.html)

# [Assess Data Availability](assess_availability.html)

# [Quality Control](quality_control.html)

## [Rainfall](quality_control.html)

## [Temperatures](quality_control.html)

# [Homogeneity Test](homogeneity_test.html)

# [Merging Rainfall](merging_rainfall.html)

# [Merging Temperature](merging_temperature.html)

# [Scale merged data](scale_merged_data.html)

# [Updating Dekadal Rainfall](update_dekadal_rainfall.html)

# [Validation](validation.html)

# [Data Extraction](data_extraction.html)

# [Summary Statistics](summary_statistics.html)

# [Derived Climate Variables](derived_climate_variables.html)

# [Climatologies and Anomalies](climatologies_anomalies.html)

# [Spatial Analysis](spatial_analysis.html)

# [Daily Rainfall Analysis](daily_rainfall_analysis.html)

# [Rainy Season Characteristics](rainy_season_characteristics.html)

# [Rainy Season Analysis](rainy_season_analysis.html)

# [Climate Extremes Indices](climate_extremes_indices.html)

# [Drought Indices](drought_indices.html)

# [Plot CDT Station Data](plot_station_data.html)

# [Plot CDT Gridded Dataset](plot_cdt_dataset.html)

# [Plot NetCDF Data](plot_ncdf.html)

# [Plot multiple data sets](plot_multiple_dataset.html)

</div>

```{r, child = '_generated_date.Rmd'}
```
